An Accurate Model for SARS-CoV-2 pooled RT-PCR Test Errors

### 2021-03-24 comment.tex
damn this paper

\subsection{Our model}
Since the essence of the failed common assumption is that
amplification of all samples occurs only once, we assume the opposite:
amplification of viral RNA occurs for each sample independently.
According to \cite{Simplistic1, Simplistic2, Kim, OptimalDorfmanPool},
erroneous detection (a false-positive) does not depend on the number
of negative samples in a pool. We incorporate this assumption in our
model with a small modification: We assume that an erroneous
amplification can occur in \emph{any} pool. Specifically, it is
possible that correct amplifications fail and an erroneous one occurs
simultaneously. This assumption is somewhat specific for the current
application of screening for SARS-CoV-2 via RT-PCR. For example,
cross-reactivity with other coronaviruses would have violated this
assumption, but it was ruled out in \cite{KitComparison}.

@@

The difference per person is:
\begin{equation}
  1 - \Sp(1-q\Se)^N - \Se - (1-q)^N(1-\Sp - \Se)
\end{equation}

\subsection{Scheme false-negative rate}
We calculate the false-negative rate of Dorfman's pooling scheme. The
false-negative rate for a single \emph{infected} individual,
henceforth referred to as "Donald" under our model and under the
common assumption. We distinguish three types of false-negative events
when performing pooling. A \emph{single test}'s false-negative is the
event of a negative result upon testing Donald separately, i.e., in an
RT-PCR test without pooling. A \emph{pooled} false-negative occurs
when a pooled test containing Donald's sample (and other samples)
yields a negative result, i.e., the pooling fails to detect at least
one positive result. Lastly, a \emph{scheme} false-negative occurs
when an entire pooling scheme fails to identify Donald as
infected. Our first task is to calculate Dorfman's scheme
false-negative rate. Rephrasing, we wish to answer the following
question: what is the probability of not identifying Donald as
infected under a Dorfman pooling scheme?

@@

The efficiency is words...

@@

\subsection{Imlications for group testing in different prevalence settings}
Our results have practical implications for the use of
pooling schemes. As the COVID-19 pandemic progresses, the infection prevalence in various tested populations undergoes frequent changes. Hence, as our results suggest, pooling strategies for mass testing should be used with caution in populations where infection rates are low. Such mass-tested populations often include air travel passengers \cite{JTM,RobinHood}, potentially importing infections, or asymptomatic individuals, such as children, and can be crucial for controlling outbreaks \cite{MinaScience}.

\subsection{Limitations}
Lacking large amounts of experimental data, it is hard to
assess the validity of our model. The data in \cite{Salazar} does allow us to reject the 
common assumption but not much more. However, our model is as simple as possible, given that the standard assumptions are refuted.

\subsection{Ways to reduce errors}

@@

Although our application focused on Dorfman pooling, our analysis
applies to other pooling schemes. Pooling schemes
(e.g. \cite{MatrixPooling,Lion, Kim}), require some sequence of
positive pooled results to correctly identify Donald as
infected. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
If the first pool yields a positive result, it is split in two. Then
the splitting is repeated until resulting pools are negative or
individuals are tested separately. With an initial pool size of 32,
Donald will necessarily have to test positive in pools of size 32, 16,
8, 4 and 2, as well as in a single test, for the scheme to correctly
identify him as infected. Compare this to the Dorfman scheme that
requires a positive test in a pool of size $N=8$, and an additional
single positive test to identify Donald as infected. The hierarchical
pooling scheme of \cite{Lion, Kim} will necessarily yield more
false-negatives than Dorfman pooling --- there are additional places
for it to fail.

>>>>

\subsection{Our model}
Since the essence of the failed common assumption is that
amplification of all samples occurs only once, we assume the opposite:
amplification of viral RNA occurs (or fails) for each sample
independently.  According to \cite{Simplistic1, Simplistic2, Kim,
  OptimalDorfmanPool}, erroneous detection (a false-positive) does not
depend on the number of negative samples in a pool. We incorporate
this assumption in our model with a small modification: We assume that
an erroneous amplification can occur in \emph{any} pool. Specifically,
it is possible that correct amplifications fails and an erroneous one
occurs simultaneously. This assumption is somewhat specific for the
current application of screening for SARS-CoV-2 via RT-PCR. For
example, cross-reactivity with other coronaviruses would have violated
this assumption. However, cross-reactivity was ruled out in
\cite{KitComparison}.

@@

The difference is:
\begin{equation}
  \mathbb{E}^{\text{exact}}[T] - \mathbb{E}^{\text{common}}[T] = N \Big [1 - \Sp(1-q\Se)^N - \Se - (1-q)^N(1-\Sp - \Se) \Big ]
\end{equation}

\subsection{Scheme false-negative rate}
We calculate the false-negative rate for a single \emph{infected}
individual, henceforth referred to as "Donald" under our model and
under the common assumption. We distinguish three types of
false-negative events when performing pooling. A \emph{single test}'s
false-negative is the event of a negative result upon testing Donald
separately, i.e., in an RT-PCR test without pooling. A \emph{pooled}
false-negative occurs when a pooled test containing Donald's sample
(and other samples) yields a negative result, i.e., the pooling fails
to detect at least one positive result. Lastly, a \emph{scheme}
false-negative occurs when an entire pooling scheme fails to identify
Donald as infected. Our first task is to calculate Dorfman's scheme
false-negative rate. Rephrasing, we wish to answer the following
question: what is the probability of not identifying Donald as
infected under a Dorfman pooling scheme?

@@

\subsection{Scheme efficiency}
Our results suggest that the efficiency under the common assumption is
always greater than the efficiency under our exact model. This should
not come as a surprise, as the common assumption implies larger
false-negative rates and, consequently, a lower number of tests being
utilized. As can be expected, the discrepancy is largest for lower
test sensitivity rates $\Se$.

@@

\subsection{Implications for group testing in different prevalence settings}
Our results have practical implications for the use of pooling
schemes. As the COVID-19 pandemic progresses, the infection prevalence
in various tested populations undergoes frequent changes. Hence, as
our results suggest, pooling strategies for mass testing should be
used with caution in populations where infection rates are low. Such
mass-tested populations often include air travel passengers
\cite{JTM,RobinHood}, potentially importing infections, or
asymptomatic individuals, such as children, and can be crucial for
controlling outbreaks \cite{MinaScience}.

One important consequence of the efficiency charts is that
practitioners order RT-PCR reagents, they should expect \emph{lower}
scheme efficiency, and thus order more reagents.

\subsection{Limitations}
Lacking large amounts of experimental data, it is hard to assess the
validity of our model. The data in \cite{Salazar} does allow us to
reject the common assumption but not much more. Our model is as simple
as possible, given that the standard assumptions are refuted and
dilution effects are not taken into consideration.

\subsection{Ways to reduce errors}
Although our application focused on Dorfman pooling, our analysis
applies to other pooling schemes. Pooling schemes
(e.g. \cite{MatrixPooling,Lion, Kim}), require some sequence of
positive pooled results to correctly identify Donald as
infected. Consider the hierarchical pooling scheme \cite{Lion, Kim}:
If the first pool yields a positive result, it is split in two. Then
the splitting is repeated until resulting pools are negative or
individuals are tested separately. With an initial pool size of 32,
Donald will necessarily have to test positive in pools of size 32, 16,
8, 4 and 2, as well as in a single test, for the scheme to correctly
identify him as infected. Compare this to the Dorfman scheme that
requires a positive test in a pool of size $N=8$, and an additional
single positive test to identify Donald as infected. The hierarchical
pooling scheme of \cite{Lion, Kim} will necessarily yield more
false-negatives than Dorfman pooling --- there are additional places
for it to fail.

Thus, finding better pooling schemes is imperative. One scheme that
holds great promise is DOPE \cite{DOPE}. This scheme utilizes
considerably less RT-PCR tests and has lower error rates than many
competing schemes.

@@
